Enforced Invariant Confluence

### 2016-11-26 doc/counters.tex
Added 1-iconvergence.
0-isafety => 1-iconvergence => iconfluence.

\subsection{\iconfluence{} Alternatives}
\begin{todoitemize}
  \item \todo{\istrength $\centernot \implies$ \isafety?}
  \item \todo{\isafety $\implies$ \istrength?}
  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
  \item \todo{\iconfluence $\implies$ \istrength?}
  \item \todo{\istrengthstar $\implies$ \istrength?}
  \item \todo{\istrength $\implies$ \istrengthstar?}
  \item \todo{Find something between \isafety{} and \iconfluence.}
  \item \todo{1-iconfluence with derived txns to start}
\end{todoitemize}

@@

  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
    \leqnomode
    \begin{flalign*}
      & \exists G_0 \subseteq \dbs \times \dbs. & \\
      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
      & \forall c.\>
        \exists P_1, \ldots, P_n \subseteq \dbs.\>
        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
    \end{flalign*}
    \reqnomode
    where $I$ is viewed as a set of database states.
\end{itemize}

@@

@@

  \begin{tikzpicture}
    \node[] (0-isafety) at (0, 0) {0-\isafety};
    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
    \node[below=of istrengthstar] (istrength)     {\istrength};

    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
    \draw[ultra thick, impl] (0-isafety) -> (istrength);
    \draw[impl] (iconfluence) -> (k-iconfluence);
    \draw[impl] (k-iconfluence) -> (1-iconfluence);
    \draw[impl] (istrengthstar) -> (ipreservation);
    \draw[impl] (0-isafety) -> (1-isafety);
    \draw[impl] (1-isafety) -> (k-isafety);
    \draw[impl] (istrength) -> (istrengthstar);
  \end{tikzpicture}

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{s2a}{0, 1}
      \dnode{s2b}{0, 0}
      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};

      \inode[label=above:$D$]{s3a}{2, 1}
      \nnode[]{s3b}{2, 0}
      \nnode[label=above:$D'$]{s3c}{3, 1}
      \nnode[]{s3d}{3, 0}
      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
    \end{tikzpicture}
    \caption{\istrengthstar}
    \label{fig:istrengthstar}
  \end{subfigure}

  \caption{}
  \label{}
\end{figure}

>>>>

\subsection{\iconfluence{} Alternatives}
\begin{todoitemize}
  \item \todo{\istrength $\implies$ \istrengthstar?}
  \item \todo{\iconfluence $\implies$ 1-\iconvergence?}
  \item \todo{Find something between \isafety{} and \iconfluence.}
  \item \todo{1-iconfluence with derived txns to start}
\end{todoitemize}

@@

  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
    \leqnomode
    \begin{flalign*}
      & \exists G_0 \subseteq \dbs \times \dbs. & \\
      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
      & \forall c.\>
        \exists P_1, \ldots, P_n \subseteq \dbs.\>
        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
    \end{flalign*}
    \reqnomode
    where $I$ is viewed as a set of database states.

  \item \textbf{1-\iconvergence.}
    $\forall D, D', D'', c', c''.\>
      I(D) \land I(D') \land I(D'') \land
      I(D \circ \denote{c'}(D')) \land
      I(D \circ \denote{c''}(D'')) \implies
      I(D \circ \denote{c'}(D') \circ \denote{c''}(D''))$
\end{itemize}

@@

\begin{claim}\label{clm:0-isafety-implies-1-iconvergence}
  0-\isafety{} $\implies$ 1-\iconvergence{}.
\end{claim}
\begin{proof}
  Consider an arbitrary set of \imp{} transactions $T$ that is \iconfluent{}
  with respect to invariant $I$. We show that $T$ is 1-\iconvergent{} with
  respect to $I$.

  Consider arbitrary database states $D$, $D'$, $D''$ and arbitrary
  transactions $c'$, $c''$ where $I(D)$, $I(D')$, $I(D'')$, $I(D \circ
  \denote{c'}(D'))$, and $I(D \circ \denote{c''}(D''))$. Applying 0-\isafety{}
  to $D \circ \denote{c'}(D')$, $D''$, and $c''$, we have $I(D \circ
  \denote{c'}(D') \circ \denote{c''}(D''))$.
\end{proof}

\begin{claim}\label{clm:1-iconvergence-implies-iconfluence}
  1-\iconvergence{} $\implies$ \iconfluence{}.
\end{claim}
\begin{proof}
  Consider a set of \imp{} transactions $T$ that is 1-\iconvergent{} with
  respect to an invariant $I$. Consider an arbitrary database state $D$, two
  \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
  show $D \circ C^c \circ C^d$ satisfies $I$.

  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
  $j, k$ if
    $j \leq m$,
    $k \leq n$, and
    $i = j + k$, then
    $I(D \circ C^c_j \circ C^d_k)$
  %
  The base case, when $i = 0$, is trivial. $i = j = k = 0$, and $D \circ C^c_0
  \circ C^d_0 = D$ satisfies $I$ by assumption.
  %
  For the inductive step, consider arbitrary natural numbers $i, j, k$ where
    $0 < i \leq m + n$,
    $j \leq m$,
    $k \leq n$, and
    $i = j + k$.
  If $j = 0$, then $D \circ C^c_0 \circ C^d_k = D \circ C^d_k$ satisfies $I$ by
  assumption. A symmetric argument holds for $k=0$. Otherwise, $j, k > 0$. Let
  \newcommand{\wh}{\widehat}
  \begin{align*}
    \wh{D}   &\defeq D \circ C^c_{j-1} \circ C^d_{k-1}, \\
    \wh{D'}  &\defeq D \circ C^c_{j-1}, \\
    \wh{D''} &\defeq D \circ C^d_{k-1}, \\
    \wh{c'}  &\defeq c_j,\ \text{and} \\
    \wh{c''} &\defeq d_k.
  \end{align*}
  By the inductive hypothesis, we have
    $I(\wh{D})$,
    $I(\wh{D'})$,
    $I(\wh{D''})$,
    $I(\wh{D} \circ \denote{\wh{c'}}(\wh{D'}))
      = I(D \circ C^c_{j} \circ C^d_{k-1})$, and
    $I(\wh{D} \circ \denote{\wh{c''}}(\wh{D''}))
      = I(D \circ C^c_{j-1} \circ C^d_{k})$.
  Applying 1-\iconvergence{} to
    $\wh{D}$,
    $\wh{D'}$,
    $\wh{D''}$,
    $\wh{c'}$, and
    $\wh{c''}$,
  we have
    $I(\wh{D} \circ \denote{\wh{c'}}(\wh{D'})
              \circ \denote{\wh{c''}}(\wh{D''}))
      = I(D \circ C^c_j \circ C^d_k)$.
\end{proof}

@@

  \begin{tikzpicture}
    \node[] (0-isafety) at (0, 0) {0-\isafety};
    \node[below=of 0-isafety]     (1-isafety)      {1-\isafety};
    \node[below=of 1-isafety]     (k-isafety)      {$k$-\isafety};
    \node[right=of 0-isafety]     (istrengthstar)  {\istrengthstar};
    \node[above=of istrengthstar] (ipreservation)  {\ipreservation};
    \node[right=of istrengthstar] (iconfluence)    {\iconfluence};
    \node[below=of iconfluence]   (k-iconfluence)  {$k$-\iconfluence};
    \node[below=of k-iconfluence] (1-iconfluence)  {1-\iconfluence};
    \node[below=of istrengthstar] (istrength)      {\istrength};
    \node[below=of istrength]     (1-iconvergence) {1-\iconvergence};

    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
    \draw[ultra thick, impl] (0-isafety) -> (istrength);
    \draw[impl] (iconfluence) -> (k-iconfluence);
    \draw[impl] (k-iconfluence) -> (1-iconfluence);
    \draw[impl] (istrengthstar) -> (ipreservation);
    \draw[impl] (0-isafety) -> (1-isafety);
    \draw[impl] (1-isafety) -> (k-isafety);
    \draw[impl] (istrength) -> (istrengthstar);
    \draw[impl] (0-isafety) -> (1-iconvergence);
    \draw[impl] (1-iconvergence) -> (iconfluence);
  \end{tikzpicture}

@@

  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{s2a}{0, 1}
      \dnode{s2b}{0, 0}
      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};

      \inode[label=above:$D$]{s3a}{1, 1}
      \nnode[]{s3b}{1, 0}
      \nnode[label=above:$D'$]{s3c}{2, 1}
      \nnode[]{s3d}{2, 0}
      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
    \end{tikzpicture}
    \caption{\istrengthstar}
    \label{fig:istrengthstar}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \begin{tikzpicture}[scale=2]
      \inode[label=above:$D$]{D}{0, 0}
      \inode[]{l1}{$(D) + (240:1)$}
      \inode[]{r1}{$(D) + (300:1)$}
      \dnode{m}{$(l1) + (300:1)$}

      \inode[label=above:$D'$]{D'}{$(r1) + (1.5, 0.5)$}
      \nnode[]{x}{$(D') + (0, -1)$}

      \inode[label=above:$D''$]{D''}{$(D') + (1, 0)$}
      \nnode[]{y}{$(D'') + (0, -1)$}

      \draw[txn] (D) -- (l1) node[midway, left] {$\denote{c'}(D')$};
      \draw[txn] (D) -- (r1) node[midway, right] {$\denote{c''}(D'')$};
      \draw[txn] (l1) -- (m) node[midway, left] {$\denote{c''}(D'')$};
      \draw[txn] (r1) -- (m) node[midway, right] {$\denote{c'}(D')$};

      \draw[txn] (D') -- (x) node [midway, left] {$\denote{c'}(D')$};

      \draw[txn] (D'') -- (y) node [midway, left] {$\denote{c''}(D'')$};
    \end{tikzpicture}
    \caption{1-\iconvergence}
    \label{fig:1-iconvergence}
  \end{subfigure}

  \caption{
    An illustration of \iconfluence{} alternatives. \todo{Describe drawings.}
  }
  \label{fig:iconfluence-alternatives}
\end{figure}
